\textbf{\sc Towards Deliberative Control of Autonomous Robots}

### 2012-01-24 arch.tex
Added som details about deliberation. Still need a lot of refinement.

In T-REX architecture reactors have two core functions;
synchronization where the main focus is to track and refine the
evolution of the state of the world at any single tick, and
deliberation which focus on providing a plan to the agent in order to
complete the reactor {\em Internal} goals. These two processes are
intertwined in the sense that one influences the outcome of the
other.

@@

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=0.5\columnwidth]{figs/synch-relation}
  \caption{Ilustration of the synchronization flaws in a reactor. The
    reactor receive new observations when they are produced by the
    owner(s) of its internal timelines. The line after the last token
    of each timeline represent the domain of possible values for the
    end of this token. At every tick $\tau$ the reactor needs to
    integrate the {\em External} state information it received and --
    based on its model $\mathcal{M}$ -- resolve its {\em Internal}
    state that will then be provided by the architecture to other
    reactors using these state variables.}
  \label{fig:synch:flaw}
\end{figure}

@@

@@

>>>>

In T-REX architecture reactors have two core functions;
synchronization where the main focus is to track and refine the
evolution of the state of the world at any single tick, and
deliberation which focus on providing a plan to the agent in order to
complete the reactor {\em Internal} goals. These two processes are
intertwined in the sense that one influences the outcome of the
other. They are represented as two europa solvers modifying the same
plan database and for which the execution is managed by the T-REX 
framework as time advance during the execution of the system:
\begin{enumerate}
\item The synchronization solver is a specialized europa solver that
  will integrate in the plan new information about the evolution of
  the External state variable and ensure that the reactor propagate
  these in order to both identify its current state and inform
  other reactors of any state change on its internal timelines. This
  solver is summoned at the beginning of every single tick.
\item The deliberation solver is managing the deliberation process of
  the reactor in order to produce a new plan or alter its current plan
  as new goals are given to the reactor or the synchronization solver
  identifying a conflict between the current state of the world and
  the expectations of the previous plan. This process can  span other
  multiple tick and therefore can be interrupted at any single tick in
  order for the synchronization solver to do its task.
\end{enumerate}

While synchronization is assumed to be fast enough that it can be
resolved in less than a tick. Deliberation on the other hand can take
several ticks reflecting that automated planning can take time
depending on how far in the future the planner is looking ahead and
the complexity of the model. T-REX architecture, and more specifically
the europa based reactor in our case, takes this into account by
considering synchronization and deliberation as two concurrent tasks
with their own timing requirements and priority. 
While the two processes are separate, they share the same plan
internal to a reactor and therefore at every synchronization cycle the
planning process is informed of the new world state and its impact on
the plan. Conversely, when the planning process eventually finds a
solution, synchronization is informed about generating the planned
state values on the {\em external} state variables. This {\em
  external} plan defines the set of {\em goals} for this state
variable managed by another {\em reactor}. Fig. \ref{fig:tick-exec}
shows a conceptual view of these two tasks.

\begin{figure}[!htbp]
  \centering
  \vskip-1pc
  \includegraphics[width=0.55\columnwidth]{figs/tick-cycle}
  \caption{\small \texttt{T-REX} reactor execution cycle: {\em
      Deliberation} is interrupted by {\em synchronization} at the
    beginning of every {\em tick} allowing integration of state
    information.}
  \label{fig:tick-exec}
  \vskip-0.8pc
\end{figure}

@@

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=0.5\columnwidth]{figs/synch-relation}
  \caption{Illustration of the synchronization flaws in a reactor. The
    reactor receive new observations when they are produced by the
    owner(s) of its internal timelines. The line after the last token
    of each timeline represent the domain of possible values for the
    end of this token. At every tick $\tau$ the reactor needs to
    integrate the {\em External} state information it received and --
    based on its model $\mathcal{M}$ -- resolve its {\em Internal}
    state that will then be provided by the architecture to other
    reactors using these state variables.}
  \label{fig:synch:flaw}
\end{figure}

@@

Synchronization remains here a deliberation process, and for this
reason there's always the risk that the search of a solution is not a
straightforward sequence of unit decision. As such process is required
by the architecture at every ticks for every reactors presently
running this need to be carefully designed. At the engine level we
made careful choices in term of both the focus of the search -- limited
only to tokens that can potentially overlap the current tick -- and
its heuristic where the internal state flaw is considered last during
the search. The later proved to limit the amount of backtrack in the
search for a solution -- or at least how deep we need to backtrack in
our decision tree -- similarly the sequence in which we evaluate the
possible resolution for such a flaw -- presented previously -- was
selected according to simple assumptions that takes benefit of a
possibly already existing plan by assuming that if the outcome of
synchronization appears consistent with our current plan then it
probably reflects that this reactor plan is executed. Such assumption
helps direct the internal state evaluation and, in nominal cases (where
we assume that reactors are acting in concert), proved to be a sound
solution.

One final aspect to keep into account is that in T-REX point of view
synchronization is the only critical task of a reactor. Indeed, such
process is used in order for the agent to identify its
current state and maintain it consistent throughout the agent
life-time. A reactor not being able to find a consistent solution
during synchronization -- meaning that after evaluating all the
possible solutions provided by its model it did not found any plan
that is not inconsistent -- not only expresses that this reactor is
not able anymore to explain the state of the world but may jeopardize
other reactors that relies on its {\em internal} state. For this
reason a failure to do synchronization immediately result on the
agent killing this reactor and notifying any reactor that depends on
its {\em internal} state variables by posting the special observation
\textsf{Failed}. By doing so we allow for graceful degradation where
other reactors can readapt their plan as they receive this
\textsf{Failed} observation.

@@

The planning process inside a reactor is directly based on the way
europa work which was described in previous section. Still some simple
alterations were made in  order to both include deliberation
information one reactor provide to the architecture -- such as its
deliberation latency and its planning look-ahead both expressed in 
ticks -- and the fact that we potentially need to interleave this
deliberation with not the synchronization process but also with 
the deliberation of other reactors\footnote{Current implementation of
  T-REX is running on a single process and it is the responsibility of
  the agent itself to emulate the multi-threading of the reactors
  deliberation and synchronization}. For this reason we 
decompose the algorithm \ref{alg:europa:solve} inside the reactor by
having the recursive calls decomposed in single steps. This step
function will be called after synchronization and as long as the
previous call in this tick did not return \textsf{true} -- which would
express that the current solution is complete.
{\em will need to change that in algorithm environment and refine it}

\begin{verbatim}
bool step(PartialPlan plan, DecisionStack stack) 
   if( isInconsistent(plan) ) {
     relaxPlan(plan, stack);
     notify("No plan");
     return true;
   } 
   Flaw flaw = chooseFlaw(flaw);
   if( flaw==NULL )
     return true;
   DecisionPoint decision = makeDecisionPoint(flaw, plan);
   while(1) {
     while( decision.hasNext() ) {
       PartialPlan pp = decision.executeNext();
       if( isInconsistent(pp) ) {
         decision.undo();
       }  else {
          plan = pp;
          stack.push(decision);
          return chooseFlaw(flaw)==NULL;
       }
     }
     if( decisionStack.empty() ) {
       notify("No plan");
       return true;
     } else {
        repeat {
          decision = decisionStack.pop();
          decision.undo();
        } until( decision.hasNext() );
        return false;
     }
   } 
}
\end{verbatim}

One critical aspect of this deliberation step is to ensure that at the
exit of this call the $plan$ is not proven inconsistent. Indeed, it is
always possible that the next call made by the agent is the
synchronization which would immediately fail if the $plan$ is
inconsistent before any solving attempt made for synchronization. 
For this reason if the plan is proven inconsistent during a
deliberation step we either backtrack in the decision stack until we
find a node fo the tree that have remaining alternate solution or
simply fully relax the plan in situation there's absolutely no other
alternative  (by removing all the goals and keeping
only recent observations produced). These constraints are yet again
related to the fact that for the agent the only crucial functionality
of a reactor is to resolve successfully its synchronization.

The number of flaw between two steps can evolve for the following
reasons that are exogenous to deliberation :
\begin{enumerate}
\item A new goal has been posted on one of the reactor timeline. This
  is usually due to another reactor finding a plan and posting the
  corresponding new objective to this reactor.
\item The outcome of synchronization created new flaws that did not
  need to be resolved during synchronization.
\item A previously requested goal has be recalled by the initial
  requester. This can happen when the reactor did identify that is
  initial plan is no longer valid. For example as synchronization
  proved this plan to be inconsistent.
\end{enumerate}
These 3 events will potentially produce new flaws to be resolved while
any single deliberation step will usually reduce the number of flaws
to be resolved.

When the number of flaw is 0 A plan solution is found. This has two
effects:
\begin{enumerate}
\item This reactor does not need to deliberate until the next synchronization.
\item The {\em external} part of this plan can then be posted which
  eventually (at the end of the current tick if these goals overlap
  the owner planning window) will be dispatched as goals triggering
  new deliberation on these reactors.
\end{enumerate}

### 2012-01-24 europa.tex
Added som details about deliberation. Still need a lot of refinement.

	A constraint programming problem consists of a set of variables V={x1,..,xN}, where each variable takes values from a domain d1,..,dN. Domains are most often discrete and finite, but there are numerous CP implementations associated with continuous domains. Given a defined conjunctive set of constraints on the variables: C={c1(x1,..,xN), ..., cK(x1,..,xN)}, the objective is to find one or more value assignments to V where all constraints are satisfied.

@@

- The Constraint Reasoning Engine (CRE): manages variables, the domains from which they can take values, and constraints that define relationships among variables. It also provides an efficient arc consistency mechanism (TODO: ref to AC-3). The CRE is designed so that specialized reasoning algorithms for specific constraints can be easily and efficiently plugged in.

- The Plan Database (PDB): manages object types and objects, which are a mechanism to group variables to more naturally model the real world in much the same way that object oriented programming does. Also manages tokens, which are a mechanism to group variables to represent temporally scoped state. Objects and tokens can be used to model planning domains in a much more natural and extensible way that a pure CSP approach can (TODO: provide snippet?)

- The Rules Engine (RE): manages causal dependencies in a planning model (TODO: explain)

- The language modules: NDDL, ANML, and other higher level modeling languages can me used to create domain models and problem instances (TODO: point advantages over pure programming API approaches, provide snippets)

- API Layer: makes all the services available so that EUROPA can be used to build applications.

- Extension Modules bundled with EUROPA: Temporal Network, Resources, NDDL, ANML.

1.3 Modeling

@@

// Function to solve a partial plan by successive refinement
// until all flaws are resolved.
bool solve(PartialPlan plan)
{
	// Propagate the constraints to test for inconsistency. If found to be
	// inconsistent, then we can return false since this is a dead-end i.e. no refinements
	// to p can yield a consistent plan.

	if(isInconsistent(plan))
		return false;

	// Non-deterministically choose a flaw from the set of available flaws.
	Flaw flaw = chooseFlaw(plan);

	// If there are no flaws, then p is complete and we can terminate with success.
	if(flaw == NULL)
		return true;

	// Otherwise we formulate a decision point which is a branch in the search space. Each
	// choice is a particular refinement operation and the DecisionPoint collects all possible
	// refinement operations for the given flaw.
	DecisionPoint decision = makeDecisionPoint(flaw, plan);

	// Continue as long as we have something to try
	while(decision.hasNext()) {
		// A new partial plan is obtained by application of a refinement operator. Note that the ordering
    		// over refinement operators to select is a non-deterministic step.
		PartialPlan pp = decision.executeNext();

		// Recursive call to solve the new planning problem. If successful, then we are done.
		if(solve(pp))
			return true;
		else // Otherwise, retract the last refinement operation
			d.undo();
  	}

	// If we arrive here, then we have exhausted all options to resolve the flaw, including the case where
  	// no options were available initially. Thus the problem cannot be solved.
  	return false;
}

The algorithm takes as input a partial plan p and returns true if a complete and consistent refinement of p could be found (or if p is initially complete and consistent), and false otherwise. This algorithm provides for a sound and complete search, assuming that no flaws or available refinement operators are pruned unnecessarily. Dead-ends in the search are discovered through constraint propagation. Constraint propagation is a vehicle for evaluating the consistency of a partial plan and also for filtering infeasible values from consideration prior to commitment, allowing in some cases a strong look-ahead capability which is essential for tractable search. Consistency testing is initiated by the isConsistent procedure. The algorithm permits a heuristically controlled search by applying orderings for chooseFlaw and makeDecisionPoint. It results in a chronologically-backtracking, search.

>>>>

	A constraint programming problem consists of a set of variables V={x1,..,xN}, where each variable takes values from a domain d1,..,dN. Domains are most often discrete and finite, but there are numerous CP implementations associated with continuous domains. Given a defined conjunctive set of constraints on the variables: $C=\{c1(x1,..,xN), ..., cK(x1,..,xN)\}$, the objective is to find one or more value assignments to V where all constraints are satisfied.

@@

\begin{enumerate}

	\item The Constraint Reasoning Engine (CRE): manages variables, the domains from which they can take values, and constraints that define relationships among variables. It also provides an efficient arc consistency mechanism (TODO: ref to AC-3). The CRE is designed so that specialized reasoning algorithms for specific constraints can be easily and efficiently plugged in.

	\item The Plan Database (PDB): manages object types and objects, which are a mechanism to group variables to more naturally model the real world in much the same way that object oriented programming does. Also manages tokens, which are a mechanism to group variables to represent temporally scoped state. Objects and tokens can be used to model planning domains in a much more natural and extensible way that a pure CSP approach can (TODO: provide snippet?)

	\item The Rules Engine (RE): manages causal dependencies in a planning model (TODO: explain)

	\item The language modules: NDDL, ANML, and other higher level modeling languages can me used to create domain models and problem instances (TODO: point advantages over pure programming API approaches, provide snippets)

	\item API Layer: makes all the services available so that EUROPA can be used to build applications.

	\item Extension Modules bundled with EUROPA: Temporal Network, Resources, NDDL, ANML.

\end{enumerate}

\subsection{Modeling}
\label{sec:europa:modeling}

@@

\begin{verbatim}

01.  bool solve(PartialPlan plan)
02.   {
03.      if (isInconsistent(plan))  return false;
04.
05.      Flaw flaw = chooseFlaw(plan);
06.      if (flaw == NULL) return true;
07.
08.      DecisionPoint decision = makeDecisionPoint(flaw, plan);
09.
10.      while(decision.hasNext()) {
11.          PartialPlan pp = decision.executeNext();
12.
13.          if (solve(pp)) return true;
14.          else d.undo();
25.      }
16.
17.      return false;
18.  }

\end{verbatim}

The algorithm takes as input a partial plan p and returns true if a complete and consistent refinement of p could be found (or if p is initially complete and consistent), and false otherwise.

\begin{enumerate}
	\item Line 3: Propagate the constraints to test for inconsistency. If found to be inconsistent, then we can return false since this is a dead-end i.e. no refinements to p can yield a consistent plan.
	\item Line 5: Choose a flaw from the set of available flaws.
	\item Line 6: If there are no flaws, then p is complete and we can terminate with success.
	\item Line 8: Otherwise we formulate a decision point which is a branch in the search space. Each choice is a particular refinement operation and the DecisionPoint collects all possible refinement operations for the given flaw.
	\item Line 10: We keep trying until the chosen flaw is resolved, or until we run out of resolvers to try.
	\item Line 11: A new partial plan is obtained by application of a refinement operator. Note that the ordering over refinement operators to select is a non-deterministic step.
	\item Line 13: Recursive call to solve the new planning problem. If successful, then we are done.
	\item Line 14: Otherwise, retract the last refinement operation and move on to try the next one.
	\item Line 17:  If we arrive here, then we have exhausted all options to resolve the flaw, including the case where no options were available initially. Thus the problem cannot be solved.
\end{enumerate}

This algorithm provides for a sound and complete search, assuming that no flaws or available refinement operators are pruned unnecessarily. Dead-ends in the search are discovered through constraint propagation. Constraint propagation is a vehicle for evaluating the consistency of a partial plan and also for filtering infeasible values from consideration prior to commitment, allowing in some cases a strong look-ahead capability which is essential for tractable search. Consistency testing is initiated by the isConsistent procedure. The algorithm permits a heuristically controlled search by applying orderings for chooseFlaw and makeDecisionPoint. It results in a chronologically-backtracking, search.
